---
title: "Medication data suggest adverse impact of COVID pandemic on cardiovascular disease prevention: A descriptive and interrupted-time series analysis"
date: "Date compiled: `r format(Sys.time(), '%a %d %B %Y')`"
toc-title: Contents
always_allow_html: true
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
    code_folding: hide
---

```{r setup, include=FALSE}
pkgs <- c(
  "assertr",
  "beepr",
  "broom",
  "dplyr",
  "dtplyr",
  "forcats",
  "ggplot2",
  "ggthemes",
  "knitr",
  "kableExtra",
  "mice",
  "janitor",
  "lubridate",
  "qs",
  "rmarkdown",
  "sailr",
  "scales",
  "stringr",
  "readr",
  "survival",
  "tableone",
  "tidyr",
  # Fatemeh uses the following:
  "RODBC",
  "Cairo",
  "lattice",
  "getopt",
  "gtsummary"
)

for (pkg in pkgs) {
  suppressWarnings(
    suppressPackageStartupMessages(
      library(pkg, character.only = TRUE)
    )
  )
  message("\t", pkg, sep = "")
}
```

```{r connect, include=FALSE}

source("S:/0000 - Analysts Shared Resources/r-share/login_box.r");
login = getLogin();
#Connect and get the database
sql = odbcConnect('PR_SAIL',login[1],login[2]);
login = 0 # this will make your password anonymous

```


```{r timerange, echo=FALSE, message=FALSE}
study_start <- ymd("2018-03-01")
study_end   <- ymd("2021-08-31")

```

### Preamble

#### IGRP project WMCC-CCU014

#### Generated by

 - fatemeh.torabi@swansea.ac.uk
 - a.akbari@swansea.ac.uk
 

## Introduction

This work is an implementation of Dale et al.'s work in SAIL Databank. Main aim is to analyse COVID-19 effect on prescribtion and dispensing of cardiovascular related medication. The current version of the outputs only contains the dispensing branch of the work pending revision and mapping of READ codes to enable retrival of prescribing data from Welsh Longitudinal General Practice dataset.

### Figure 1

Flowchart showing selection of analytical datasets from WDDS : Dispensing branch 

```{r Fig1, echo = FALSE, message = FALSE}

Fig1.DISP <- 
sqlQuery(sql,"
SELECT * FROM (
	SELECT 
	'1' AS Row , 
	'Wales: WDDS All Dispensed medication' AS Description , 
	count(*) medication_count FROM SAILWWMCCV.CCU014_COHORT_DISPENSED

UNION 

	SELECT 
	'2' AS Row,
	'Wales: Valid ID & WOB; alive on 2018-01-04, gndr M or F'  AS description, 
	count(*) medication_count 
	FROM SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')

UNION 
	SELECT 
	'3' AS Row,
	'Wales: BNF Chapters 1-15'  AS description, 
	count(*) medication_count 
	FROM SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')
UNION 
	SELECT 
	'4' AS Row,
	'Wales: LOSA=W'  AS description, 
	count(*) medication_count 
	FROM SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
UNION 
	SELECT 
	'5' AS Row,
	'Wales: >=18 & <112 years old'  AS description, 
	count(*) medication_count 
	FROM SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
UNION 	
	SELECT 
	'6' AS Row,
	'Wales: CVD medication'  AS description, 
	count(*) medication_count 
	FROM SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
UNION 	
	SELECT 
	'7' AS Row,
	'Wales: incident medication'  AS description, 
	count(*) medication_count 
	FROM  SAILWWMCCV.CCU014_COHORT_DISPENSED_rows
	WHERE dt_prescribed >= '2019-03-01' and drug_seq=1 AND KEY_GROUP IS NOT NULL
)ORDER BY ROW;
         ")

Fig1.DISP %>% 
kable(
    digits = 1,
    format.args = list(big.mark = ","),
    col.names = c(
        "No.",
        "Description",
        "Dispensed medication count"
    )
)
```

### Table 1 

Rates and rate differences of CVD medicines dispensed by year (1st March to end February)' all CVD medicines and split by key CVD sub-groups

```{r Table1, echo = FALSE, message = FALSE}

Tab1.S1 <- sqlQuery(sql,
                    "

SELECT * FROM (
	SELECT DISTINCT 
	'a-0' AS ROW, 
	'All CVD medicines' AS description, NULL AS Number_of_dispensed_items, NULL AS Patient_population
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED

UNION 
	SELECT 
	'b-1' AS ROW, 
	'March 2018-2019' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2018-03-01' AND '2019-02-28'
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
UNION 
	SELECT 
	'c-2' AS ROW, 
	'March 2019-2020' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2019-03-01' AND '2020-02-29'
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 

UNION 
	SELECT 
	'd-3' AS ROW, 
	'March 2020-2021' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2020-03-01' AND '2021-02-28'
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
	
UNION	
---4 CVD MEDICINES
	SELECT DISTINCT 
	'e-4' AS ROW, 
	'4 CVD medicines - antihypertensive, lipid-lowering, type-2 diabetes, insulin' AS description, NULL AS Number_of_dispensed_items, NULL AS Patient_population
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	
UNION 
	SELECT 
	'f-5' AS ROW, 
	'March 2018-2019' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2018-03-01' AND '2019-02-28'
	AND key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 

UNION 
	SELECT 
	'g-6' AS ROW, 
	'March 2019-2020' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2019-03-01' AND '2020-02-29'
	AND key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 

UNION 
	SELECT 
	'h-7' AS ROW, 
	'March 2020-2021' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2020-03-01' AND '2021-02-28'	
	AND key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 

UNION
---7 CVD MEDICINES
	SELECT DISTINCT 
	'i-8' AS ROW, 
	'7 CVD medicines - AF, angina, DOACs,warfarins,heparins,antiplatelets,heart_failure' AS description, NULL AS Number_of_dispensed_items, NULL AS Patient_population
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED

UNION 
	SELECT 
	'j-9' AS ROW, 
	'March 2018-2019' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2018-03-01' AND '2019-02-28'
	AND key_group IN ('AF','angina','anticoagulant_DOAC','anticoagulant_warfarins','anticoagulants_heparins','antiplatelets_secondary','heart_failure')	
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 

UNION 
	SELECT 
	'k-10' AS ROW, 
	'March 2019-2020' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2019-03-01' AND '2020-02-29'
	AND key_group IN ('AF','angina','anticoagulant_DOAC','anticoagulant_warfarins','anticoagulants_heparins','antiplatelets_secondary','heart_failure')	
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 

UNION 
	SELECT 
	'l-11' AS ROW, 
	'March 2020-2021' AS description, count(*) AS cnt, count(DISTINCT alf_e) AS ALF
	FROM 	
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE dt_prescribed BETWEEN '2020-03-01' AND '2021-02-28'	
	AND key_group IN ('AF','angina','anticoagulant_DOAC','anticoagulant_warfarins','anticoagulants_heparins','antiplatelets_secondary','heart_failure')	
	AND 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
	

) ORDER BY row;
                    ")



Tab1.S1 %>% 
  mutate(DISP_RATE=NUMBER_OF_DISPENSED_ITEMS/PATIENT_POPULATION) %>% 
kable(
    digits = 1,
    format.args = list(big.mark = ","),
    col.names = c(
        "No.",
        "Description",
        "Dispensed medication count",
        "Patient population",
        "Dispensed medication rate per person"
    )
)
```


### Figure 2A -
Trends in dispensed 4-CVD medicines ('insulin','antihypertensives','lipid_lowering','type_2_diabetes') over the courtse of the pandemic. 

```{r Figure2, echo = FALSE, message = FALSE}

Fig2 <- sqlQuery(sql, "
	SELECT DISTINCT 
	substring(dt_prescribed,1,7) yr_mnth,
	count(*) AS counts
	FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
	AND
  key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
  GROUP BY 
	substring(dt_prescribed,1,7);
")



library(ggplot2)
library(tidyr)


Fig2 %>%
  arrange(Fig2$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=1))+
  geom_line(color="#256fa7", alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))+
  scale_color_brewer(palette = "Paired")
#  ylim(c(0,3000000))


library(DT)
DT::datatable(Fig2, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))
```

### Figure 2B -
Trends in dispensed 7-CVD medicines ('AF','angina','anticoagulant_DOAC','anticoagulant_warfarins','anticoagulants_heparins','antiplatelets_secondary','heart_failure') over the courtse of the pandemic. 

```{r Figure2b, echo = FALSE, message = FALSE}

Fig2b <- sqlQuery(sql, "
	SELECT DISTINCT 
	substring(dt_prescribed,1,7) yr_mnth,
	count(*) AS counts
	FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
		AND key_group IN ('AF','angina','anticoagulant_DOAC','anticoagulant_warfarins','anticoagulants_heparins','antiplatelets_secondary','heart_failure')	
  GROUP BY 
	substring(dt_prescribed,1,7);
")



library(ggplot2)
library(tidyr)


Fig2b %>%
  arrange(Fig2b$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=1))+
  geom_line(color="#256fa7", alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))+
  scale_color_brewer(palette = "Paired")
#  ylim(c(0,3000000))


library(DT)
DT::datatable(Fig2b, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))
```


### Figure 3 - 
Trends in dispensed CVD medicines over course of pandemic by CVD/CVD risk factor sub-group - counts by month and year

```{r Figure3, echo = FALSE, message = FALSE}

Fig3 <- sqlQuery(sql, "
	SELECT DISTINCT 
	key_group, 
	substring(dt_prescribed,1,7) yr_mnth,
	count(*) AS counts
	FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
 key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
	GROUP BY 
	key_group,
	substring(dt_prescribed,1,7);
")



library(ggplot2)
library(tidyr)


Fig3 %>%
  arrange(Fig3$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=KEY_GROUP, colour=KEY_GROUP))+
  geom_line(alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))
# scale_color_brewer(palette = "Paired")+
#  ylim(c(0,3000000))


library(DT)
DT::datatable(Fig3, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```

### Figure 4 - 

Figure 4a : Trends in incident medications dispensed by CVD/CVD risk factor sub-group - counts by month and year (4 CVD medicines - antihypertensive, lipid-lowering, type-2 diabetes, insulin)

```{r Figure4a, echo = FALSE, message = FALSE}

Fig4 <- sqlQuery(sql, "SELECT DISTINCT 
	key_group, 
	substring(dt_prescribed,1,7) yr_mnth,
	count(*) AS counts
	FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED_rows
	WHERE 
	dt_prescribed >= '2019-03-01' 
	and 
	drug_seq=1	
	AND 
	key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
	GROUP BY 
	key_group,
	substring(dt_prescribed,1,7);
	")



library(ggplot2)

Fig4 %>%
  arrange(Fig4$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=KEY_GROUP, colour=KEY_GROUP))+
  geom_line(alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))
# scale_color_brewer(palette = "Paired")+
#  ylim(c(0,3000000))


library(DT)
DT::datatable(Fig4, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```


Figure 4b : Trends in incident medications dispensed by CVD/CVD risk factor sub-group - counts by month and year (7 CVD medicines - AF, angina, DOACs,warfarins,heparins,antiplatelets,heart_failure)

```{r Figure4b, echo = FALSE, message = FALSE}

Fig4b <- sqlQuery(sql, "SELECT DISTINCT 
	key_group, 
	substring(dt_prescribed,1,7) yr_mnth,
	count(*) AS counts
	FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED_rows
	WHERE 
	dt_prescribed >= '2019-03-01' 
	and 
	drug_seq=1	
	AND 
  key_group IN ('AF','angina','anticoagulant_DOAC','anticoagulant_warfarins','anticoagulants_heparins','antiplatelets_secondary','heart_failure')	
  GROUP BY 
	key_group,
	substring(dt_prescribed,1,7);
	")



library(ggplot2)

Fig4b %>%
  arrange(Fig4b$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=KEY_GROUP, colour=KEY_GROUP))+
  geom_line(alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))
# scale_color_brewer(palette = "Paired")+
#  ylim(c(0,3000000))


library(DT)
DT::datatable(Fig4b, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```

### Figure 5 - 
Trends in monthly counts of dispensed CVD medicines () stratified by: Region, age, sex, ethnicity

#### Figure 5a - By age band

```{r Figure5a, echo = FALSE, message = FALSE}

Fig5a <- sqlQuery(sql, "
		SELECT DISTINCT 
	substring(dt_prescribed,1,7) YR_MNTH,
	CASE
	WHEN dispensed_age BETWEEN '18' AND '29' THEN '18-29'
	WHEN dispensed_age BETWEEN '30' AND '39' THEN '30-39'
	WHEN dispensed_age BETWEEN '40' AND '49' THEN '40-49'
	WHEN dispensed_age BETWEEN '50' AND '59' THEN '50-59'
	WHEN dispensed_age BETWEEN '60' AND '69' THEN '60-69'
	WHEN dispensed_age BETWEEN '70' AND '79' THEN '70-79'
	WHEN dispensed_age BETWEEN '80' AND '89' THEN '80-89'
	WHEN dispensed_age >= 90 THEN '90+' ELSE NULL END AS age_band, 
	count(*) AS counts
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
GROUP BY 
	substring(dt_prescribed,1,7),
	case
	WHEN dispensed_age BETWEEN '18' AND '29' THEN '18-29'
	WHEN dispensed_age BETWEEN '30' AND '39' THEN '30-39'
	WHEN dispensed_age BETWEEN '40' AND '49' THEN '40-49'
	WHEN dispensed_age BETWEEN '50' AND '59' THEN '50-59'
	WHEN dispensed_age BETWEEN '60' AND '69' THEN '60-69'
	WHEN dispensed_age BETWEEN '70' AND '79' THEN '70-79'
	WHEN dispensed_age BETWEEN '80' AND '89' THEN '80-89'
	WHEN dispensed_age >= 90 THEN '90+' ELSE NULL END;
")

library(ggplot2)
library(tidyr)

Fig5a %>%
  arrange(Fig5a$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=AGE_BAND, colour=AGE_BAND))+
  geom_line(alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))


library(DT)
DT::datatable(Fig5a, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```

#### Figure 5b - By Sex 

```{r Figure5b, echo = FALSE, message = FALSE}

Fig5b <- sqlQuery(sql, "
		SELECT DISTINCT 
	substring(dt_prescribed,1,7) YR_MNTH,
	CASE 
	WHEN gndr_cd=1 THEN 'male'
	WHEN gndr_cd=2 THEN 'female' ELSE NULL END AS sex, 
	count(*) AS counts
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
GROUP BY 
	substring(dt_prescribed,1,7),
	CASE 
	WHEN gndr_cd=1 THEN 'male'
	WHEN gndr_cd=2 THEN 'female' ELSE NULL END;
")

library(ggplot2)
library(tidyr)

Fig5b %>%
  arrange(Fig5b$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=SEX, colour=SEX))+
  geom_line(alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))


library(DT)
DT::datatable(Fig5b, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```

### Figure 5c - By Region 
Wales as one region so it matches to Figure 2 above: for the 4 CVD medications

```{r Figure5c, echo = FALSE, message = FALSE}

Fig5c <- sqlQuery(sql, "
	SELECT DISTINCT 
	substring(dt_prescribed,1,7) yr_mnth,
	count(*) AS counts
	FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
	GROUP BY 
	substring(dt_prescribed,1,7);
")



library(ggplot2)
library(tidyr)


Fig5c %>%
  arrange(Fig2$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=1))+
  geom_line(color="#256fa7", alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))+
  scale_color_brewer(palette = "Paired")
#  ylim(c(0,3000000))


library(DT)
DT::datatable(Fig5c, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```


#### Figure 5d - By ethnicity

```{r Figure5d, echo = FALSE, message = FALSE}

Fig5d <- sqlQuery(sql, "
	SELECT DISTINCT 
	substring(dt_prescribed,1,7) YR_MNTH,
	CASE 
	WHEN ETHNICITY_ONS_DESC IS NOT NULL THEN ETHNICITY_ONS_DESC
	ELSE 'UNKNOWN' END AS ethnicity, 
	count(*) AS counts
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IN ('insulin','antihypertensives','lipid_lowering','type_2_diabetes')
GROUP BY 
	substring(dt_prescribed,1,7),
	CASE 
	WHEN ETHNICITY_ONS_DESC IS NOT NULL THEN ETHNICITY_ONS_DESC
	ELSE 'UNKNOWN' END;
")

library(ggplot2)
library(tidyr)

Fig5d %>%
  arrange(Fig5d$YR_MNTH)%>%
  ggplot(aes(YR_MNTH,as.numeric(COUNTS), group=ETHNICITY, colour=ETHNICITY))+
  geom_line(alpha=0.6, size=1)+
  geom_point()+
  theme_excel_new()+
  theme(axis.text.x=element_text(angle=45, size=7))


library(DT)
DT::datatable(Fig5d, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```

#### Table A - This table contains total count of prescribed medicines, total number of patients, average quantity per each prescribed pacl and total quantity for two specific drugs: 

* 0212000B0AAACAC	Atorvastatin 40mg tablets

* 0212000Y0AAADAD	Simvastatin 40mg tablets

```{r Sensitivity, message = FALSE}

STATIN <- sqlQuery(sql,"
SELECT DISTINCT 
	key_group, 
	bnf, 
	BNF_DESC ,
	year(dt_prescribed) YR,
	month(dt_prescribed) MNTH,	 
	count(*) AS counts, 
	count(DISTINCT ALF_E) AS alf_counts, 
	ROUND(AVG(QTY_PRESCRIBED_VAL),2) average_qty_prescribed,
	sum(QTY_PRESCRIBED_VAL)	total_qty_prescribed
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
	AND 
	BNF IN ('0212000B0AAACAC','0212000Y0AAADAD')
GROUP BY 
key_group, 
year(dt_prescribed),
month(dt_prescribed), 
BNF ,BNF_DESC
ORDER BY 
key_group, 
year(dt_prescribed),
month(dt_prescribed), 
BNF ,BNF_DESC; 
")

library(DT)

DT::datatable(STATIN, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))

```


## Data table: 
Extract entire data for all figures - Please note any count of <5 been removed from the extract.

```{r data, echo = FALSE, message = FALSE}

Raw <- sqlQuery(sql,"

SELECT * FROM (
	SELECT DISTINCT 
	key_group, 
	year(dt_prescribed) YR,
	month(dt_prescribed) MNTH,
	'sex' AS strat_key,
	gndr_cd AS  category, 
	count(*) AS counts
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
GROUP BY 
key_group, 
year(dt_prescribed),
month(dt_prescribed),
gndr_cd

UNION 

SELECT DISTINCT 
key_group, 
year(dt_prescribed) YR,
month(dt_prescribed) MNTH,
'Ethnicity' AS strat_key,
ETHNICITY_ONS_DESC AS  category, 
count(*) AS counts
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
GROUP BY 
key_group, 
year(dt_prescribed),
month(dt_prescribed),
ETHNICITY_ONS_DESC

UNION 

SELECT DISTINCT 
key_group, 
year(dt_prescribed) YR,
month(dt_prescribed) MNTH,
'age_band' AS strat_key,
CASE
WHEN dispensed_age BETWEEN '18' AND '29' THEN '18-29'
WHEN dispensed_age BETWEEN '30' AND '39' THEN '30-39'
WHEN dispensed_age BETWEEN '40' AND '49' THEN '40-49'
WHEN dispensed_age BETWEEN '50' AND '59' THEN '50-59'
WHEN dispensed_age BETWEEN '60' AND '69' THEN '60-69'
WHEN dispensed_age BETWEEN '70' AND '79' THEN '70-79'
WHEN dispensed_age BETWEEN '80' AND '89' THEN '80-89'
WHEN dispensed_age >= 90 THEN '90+' ELSE NULL END 
AS  category, 
count(*) AS counts
FROM 
	SAILWWMCCV.CCU014_COHORT_DISPENSED
	WHERE 
	alf_e IS NOT NULL 
	AND 
	wob IS NOT NULL 
	AND 
	(dod IS NULL OR dod > '2018-03-01')
	AND 
	gndr_cd IN ('1','2')	
	AND
	substr(bnf, 1, 2) IN ('01','02','03','04','05','06','07','08','09','10','11','12','13','14','15')	
	AND 
	SUBSTRING(LSOA2011_INCEPTION,1,1) IN ('W')
	AND 
	dispensed_age BETWEEN '18' AND '112'
	AND 
	key_group IS NOT NULL 
GROUP BY 
key_group, 
year(dt_prescribed),
month(dt_prescribed),
CASE
WHEN dispensed_age BETWEEN '18' AND '29' THEN '18-29'
WHEN dispensed_age BETWEEN '30' AND '39' THEN '30-39'
WHEN dispensed_age BETWEEN '40' AND '49' THEN '40-49'
WHEN dispensed_age BETWEEN '50' AND '59' THEN '50-59'
WHEN dispensed_age BETWEEN '60' AND '69' THEN '60-69'
WHEN dispensed_age BETWEEN '70' AND '79' THEN '70-79'
WHEN dispensed_age BETWEEN '80' AND '89' THEN '80-89'
WHEN dispensed_age >= 90 THEN '90+' ELSE NULL END 

)
where counts >= 5
ORDER BY key_group, yr, mnth, strat_key, category;
              ")

library(DT)
DT::datatable(Raw, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))
#Raw %>% 
#kable(
#    digits = 1,
#    format.args = list(big.mark = ",")
#)
```

